---
title: "Pokemon - Dragon - CP - Analysis"
author: "Alan Cheun"
date: "`r Sys.Date()`"
output: 
  word_document: 
    fig_width: 8
    fig_height: 6
header-includes:
  - \usepackage[left=0.5in, right=0.5in, top=0.5in, bottom=0.5in]{geometry}
---

```{r Librairies}
# Load necessary libraries
library(readr)
library(tidyverse)
library(corrplot)
```

```{r Dataset}

Pokemon <- read_csv("Pokemon.csv")
#View(Pokemon)
summary(Pokemon)
str(Pokemon)
```

# Task III

## Introduction

In the realm of Pokémon battles, understanding Pokémon types is paramount. Pokémon types encompass various elements or specializations that a Pokémon possesses, ranging from nature-based categories like fire, water, and ground to domain-based classifications such as fighting, ghost, and psychic. These typings play a crucial role, offering distinct advantages or disadvantages in battles against other Pokémon.

Considered the foundational principle of Pokémon battling, understanding Pokémon typings is indispensable.

The objective of this project is to conduct a regression analysis to explore the relationship between Combat Power (CP), an overall measure of the effectiveness of a Pokémon in the Pokémon game, and selected Typing from the Pokémon dataset. Combat Power is a crucial metric in Pokémon, representing the battle strength of a Pokémon. This analysis aims to identify significant factors influencing CP and provide insights into how Total stats and explore if Typing contribute to the Combat Power of Pokémon.

We will explore Dragon, considered high in rank and of formidable power, not just in Pokémon lore, across many mythologies and legends.\
We will explore Bug, considered often weak and insignificant, this will provide a sound opposite to the Dragon typing.

Finally explore Water, this is the most prevalent typing across the first 6 generations of Pokémon .

## Dataset Selection:

For this analysis, we have chosen the Pokémon dataset, which contains information about various Pokémon species, their attributes, and Combat Power. The dataset includes both categorical and numeric variables, making it suitable for exploring relationships through regression analysis. Specifically, we are interested in understanding how the total stats (Total) and whether a Pokémon typing impacts its Combat Power.

### Data Wrangling

```{r DataWrangling}
#CP function
Combat_Power <- function(attack, special_atk, defense, special_def, hp) {
  total_a <- attack + special_atk
  total_d <- defense + special_def
  numerator <- floor(total_a * (total_d^0.5) * (hp^0.5) * (0.667934^2))
  return(floor(max(10, numerator / 10)))
}
#Add Combat Power to dataset and isTyping indicator
Pokemon <- Pokemon %>%
  mutate(isDragon = ifelse(`Type 1` == "Dragon" | `Type 2` == "Dragon", 1, 0)) %>%
  mutate(isDragon = as.integer(ifelse(is.na(isDragon), 0, isDragon))) %>%
  rowwise() %>%
  mutate(isBug = ifelse(`Type 1` == "Bug" | `Type 2` == "Bug", 1, 0)) %>%
  mutate(isBug = as.integer(ifelse(is.na(isBug), 0, isBug))) %>%
  rowwise() %>%
  mutate(isWater = ifelse(`Type 1` == "Water" | `Type 2` == "Water", 1, 0)) %>%
  mutate(isWater = as.integer(ifelse(is.na(isWater), 0, isWater))) %>%
  rowwise() %>%
  mutate(CP = Combat_Power(Attack, `Sp. Atk`, Defense, `Sp. Def`, HP))

# Convert isDragon to a factor with levels "0" and "1"
Pokemon$isDragon <- as.factor(Pokemon$isDragon)
Pokemon$isBug <- as.factor(Pokemon$isBug)
Pokemon$isWater <- as.factor(Pokemon$isWater)

# Print the first few rows to check if CP values vary
head(Pokemon)

```

We are to add dummy variables that will act as a Categorical variable in the regression model analysis.

**isDragon** is whether Type 1 or Type 2 is equal to Dragon, then set to 1 for True, 0 for False.

**isBug** is whether Type 1 or Type 2 is equal to Bug, then set to 1 for True, 0 for False.

**isWater** is whether Type 1 or Type 2 is equal to Water, then set to 1 for True, 0 for False.

## Regression Model:

### Assumptions and Validation for Regression Model

```{r RegressionCorrelationCheck}
selected_columns <- c("Total", "HP", "Attack", "Defense", "Sp. Atk", "Sp. Def", "Speed", "CP")

# Create a scatterplot matrix with linear regression lines
pairs(Pokemon[selected_columns],main="Pairwise Relationships Between Pokemon Attributes")

#Correlation 
#cor(Pokemon[selected_columns])
correlation_matrix <- cor(Pokemon[selected_columns])
# Create a compact correlation plot (heatmap)
corrplot(correlation_matrix, method = "number", type = "upper", tl.cex = 0.6, main = "Correlation Matrix - Pokemon",mar = c(0, 0, 2, 2))


```

```{r RegressionAssumptionCheck}
# Perform regression analysis
cp_dragon_model <- lm(CP ~ Total + isDragon, data = Pokemon)
cp_bug_model <- lm(CP ~ Total + isBug, data = Pokemon)
cp_water_model <- lm(CP ~ Total + isWater, data = Pokemon)


par(mfrow = c(1, 2)) 

# Print regression summary for Dragon
plot(cp_dragon_model$residuals, cp_dragon_model$fitted.values, main = "Dragon - Residuals vs Fitted", xlab = "Fitted values", ylab = "Residuals")
# Plot Q-Q Residuals for Dragon model
qqnorm(cp_dragon_model$residuals,main = "Dragon - Normal Q-Q Plot")
qqline(cp_dragon_model$residuals)


# Print regression summary for Bug
plot(cp_bug_model$residuals, cp_bug_model$fitted.values, main = "Bug - Residuals vs Fitted", xlab = "Fitted values", ylab = "Residuals")
# Plot Q-Q Residuals for Bug model
qqnorm(cp_bug_model$residuals,main = "Bug - Normal Q-Q Plot")
qqline(cp_bug_model$residuals)

# Print regression summary for Water
plot(cp_water_model$residuals, cp_water_model$fitted.values, main = "Water - Residuals vs Fitted", xlab = "Fitted values", ylab = "Residuals")
# Plot Q-Q Residuals for Water model
qqnorm(cp_water_model$residuals,main = "Water - Normal Q-Q Plot")
qqline(cp_water_model$residuals)

```

In the scatterplots from "*Scatterplots for Pairwise Relationships Between* Pokémon *Attributes*", there is a pattern apparent between Total and CP, while all other pairwise plots do not have a clear pattern identified. This provides us enough evidence that the Total attribute of the Pokémon should be included in a linear regression formula for predicting CP.

\
As we look further into In the Residuals vs Fitted plot into the focused Typings, the scatter plot does not indicate any sign of a pattern, thus residuals are independent, there is equal variances.\
This is a positive sign that the requirements for linear regression model is met.\
\
In the Q-Q Residuals plot, the points are falling close along the line, thus indicating normal distribution of the residuals for us to use certain statistical methods and tests.

### Estimate the Model

```{r RegressionAnalysis}

# Print regression summary for Water
cat("LM - CP - Dragon Model")
summary(cp_dragon_model)


cat("LM - CP - Bug Model\n")
summary(cp_bug_model)


cat("LM - CP - Water Model")
summary(cp_water_model)

```

We have employed a linear regression model to investigate the associations between CP and the chosen predictors. The regression equation is given by:

#### Dragon

$$ CP = -666.06 + 3.15 \times \text{Total} + 111.97 \times \text{isDragon}$$

The p-values for the intercept, Total, and isDragon1 are all reported as '\< 2e-16', which means they are very close to zero. This suggests strong evidence to reject the null hypothesis that the corresponding coefficients are zero.

The Multiple R-squared value is 0.9014, indicating that approximately 90.14% of the variability in the response variable (CP) is explained by the linear regression model.

#### Bug

$$ CP = -677.91 + 3.20 \times \text{Total} - 15.52 \times \text{isBug}$$

The p-values for the intercept and Total are both reported as '\< 2e-16', suggesting strong evidence to reject the null hypothesis that the corresponding coefficients are zero.

The p-value for isBug1 is 0.34, which is greater than 0.05, indicating that the *coefficient for isBug1 is not statistically significant*.

#### Water

$$ CP = -684.24 + 3.20 \times \text{Total} + 13.89  \times \text{isWater}$$

The p-values for the intercept and Total are both reported as '\< 2e-16', suggesting strong evidence to reject the null hypothesis that the corresponding coefficients are zero.

The p-value for isWater1 is 0.272, which is greater than 0.05, indicating that the coefficient for *isWater1 is not statistically significant*.

```{r DragonPlot}


cb_palette <- c("red", "blue")

ggplot(Pokemon, aes(Total, CP, col = isDragon)) +
  geom_point() +
  geom_abline(aes(slope = 3.14984, intercept = -666.06293, col = "0"), size = 1.5) +
  geom_abline(aes(slope = 3.14984, intercept = -666.06293 + 111.96653, col = "1"), size = 1.5) +
  labs(title = "Combat Power vs Total - Dragon Typing Comparison",
       x = "Total",
       y = "Combat Power") +
  scale_color_manual(values = cb_palette) +
  theme_minimal()  # You can customize the theme as needed
```

```{r BugPlot}


ggplot(Pokemon, aes(Total, CP, col = isBug)) +
  geom_point() +
  geom_abline(aes(slope = 3.19637, intercept = -677.91357), col = "0") +
  geom_abline(aes(slope = 3.19637, intercept = -677.91357 + -15.51616), col = "1") +
  labs(title = "Combat Power vs Total - Bug Typing Comparison",
       x = "Total",
       y = "Combat Power") +
  scale_color_manual(values = cb_palette)
```

```{r WaterPlot}

ggplot(Pokemon, aes(Total, CP, col = isWater)) +
  geom_point() +
  geom_abline(aes(slope =  3.20268, intercept = -684.2440), col = "0") +
  geom_abline(aes(slope =  3.20268, intercept = -684.2440 + 13.88566), col = "1") +
  labs(title = "Combat Power vs Total - Water Typing Comparison",
       x = "Total",
       y = "Combat Power") +
  scale_color_manual(values = cb_palette)
```

In summary, the p-values for the Total in all models are well below 5% significance level and thus strongly support with statistical evidence that the coefficient of Total to CP is not zero.

The Multiple R-squared values is around \~0.90, indicating that approximately 90% of the variability in the response variable (CP) is explained by the linear regression model. This is consistent with expectations given the strong coefficient correlation between Total and Combat Power.

Analyzing the three regression models, the Dragon typing had a p-value that strongly suggests there is a non-zero relationship between a Pokémon being of a Dragon typing and its impact to its CP. While the other Pokémon typings used in comparison show no statistical evidence for a relationship between a Bug or Water typing to its CP. This lack of correlation for Bug and Water can be visualized by the lack of any gap between the regression lines between its true or false status in Pokémon typing.

```{r PredictCP}
#Dragon
cat("Dragon CP Prediction")
new_CP_Dragon_Pokemon <-tibble(isDragon = c("0","1","0","1","0","1","0","1","0","1","0","1","0","1"),Total = c(200,200,300,300, 400,400,500,500,600,600,700,700,800,800))

new_CP_Dragon_Pokemon$Predicted_CP <- predict(cp_dragon_model, newdata = new_CP_Dragon_Pokemon)

#Bug
#cat("Bug CP Prediction")
#predict_CP_Bug_Pokemon <-tibble(isBug = c("0","1","0","1","0","1","0","1"),Total = c(200,200, 400,400, 600,600,800,800))
#predict(cp_bug_model,predict_CP_Bug_Pokemon)
#Water
#cat("Water CP Prediction")
#predict_CP_Water_Pokemon <-tibble(isWater = c("0","1","0","1","0","1","0","1"),Total = c(200,200, 400,400, 600,600,800,800))
#predict(cp_water_model,predict_CP_Water_Pokemon)
# Plotting the ggplot with both observed and predicted values
# Convert isDragon to factor for better plotting
new_CP_Dragon_Pokemon$isDragon <- factor(new_CP_Dragon_Pokemon$isDragon)

# Plotting the ggplot with both observed and predicted values
ggplot() +
  geom_point(data = Pokemon, aes(x = Total, y = CP, col = isDragon)) +
  geom_point(data = new_CP_Dragon_Pokemon, aes(x = Total, y = Predicted_CP, shape = isDragon), col = "black", size = 3) +
  geom_abline(aes(slope = 3.14984, intercept = -666.06293, col = "0")) +
  geom_abline(aes(slope = 3.14984, intercept = -666.06293 + 111.96653, col = "1")) +
  labs(title = "Combat Power vs Total + Predicted Dragon Typing CP ",
       x = "Total",
       y = "Combat Power") +
  scale_color_manual(values = cb_palette)
```

## Conclusion:

This regression analysis contributes to our understanding of the factors influencing Combat Power in Pokémon. The highly significant Total and isDragon variables indicate their importance in predicting Combat Power and have a possibility to be deployed. The regression models in utilizing Bug and Water typing are rejected to be deployed as it failed to reject the null hypothesis that its coefficient is significant.
